Optimizing Structural Properties Of Neural Networks With Genetic Algorithms

### 2015-08-07 paper/paper.tex
paper.tex

    \newcommand{\param}[1]{\texttt{#1}}
    \newcommand{\feedback}{\param{feedback}}
    \newcommand{\gain}{\param{gain}}
    \newcommand{\size}{N}
    \newcommand{\ERprob}{p}
    \newcommand{\RINGk}{k}
    \newcommand{\ERtoRING}{r}
    \newcommand{\FFlayers}{l}

@@

The genotypes' parameters $\theta\in\Theta$ are either integer or floating
point values and can be classified as affecting (1) the topology of
$\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
$\omega\Ix{fb}$. Optimized parameters of type (1) are the
occupation probability $p$ for sparse random topologies (Erdös-Renyi
topology), the neighborhood range $k$ for ring topologies, and a ratio
parameter $q$ used to interpolate between Erdös-Renyi and ring
topologies. Optimized parameters of type (2) are the feedback strength
$\feedback\propto \omega\Ix{fb}$ and the gain value
$\gain\propto\omega\Ix{rec}$.

@@

\begin{figure}
    {\footnotesize\input{pictures/freqrange}}
    \caption{
        The covered frequency range for an unoptimized
        genotype $G_0$ as well as genotypes $G\ix{high}$ and $G\ix{low}$
        that are optimized for $C\ix{high}$ respectively $C\ix{low}$. Each
        of the three genotypes was tested against constant challenges
        $C_\nu$ that yielded simple sinusoidal waves with frequency
        $\nu$ for every seed value. The topology of all genotypes was
        a mere Erdös-Renyi topology.
        The parameters for the unoptimized genotype $G_0$ were taken to be
        default values from \cite{FORCE} (which also uses Erdös-Renyi
        topologies), while we chose genotypes $G\ix{high}$ and $G\ix{low}$
        that did particularly well when explicitly optimizing for high and
        low frequencies (see FIG. \ref{fig:freq_optimization} for more
        details).
    }
    \label{fig:freqrange}
\end{figure}

\subsection{Topology Optimizations}
To find out whether the topology can be used to improve the learning, we
performed an optimization in which the networks were an interpolation
between sparse random connections and a ring topology where
the $\RINGk$ nearest neighbors are connected (for non integer $\RINGk$, the
fractional part determines the probability to use $[\RINGk]$ or $[\RINGk]+1$).
Since we are interested in the influence of the topology, the parameters
$\gain$ and $\feedback$ are kept constant during the optimization process,
leaving the connection probability for ER topology $\ERprob$, the ring
connection length $\RINGk$ and the interpolation value $\ERtoRING$ between
ER and ring as free parameters.

The results of the optimization are depicted in FIG. \ref{}. Unfortunately,
the improvement in fitness after the first few generations in extremely
low. The development of the ration parameter, however, unambiguously shows
that the selection process prefers generators that consist mostly of ring
topology with ring parameter $\RINGk \approx 1$. The ER connection probability
shows a very broad distribution, which is simply due to the low dependence
of the generated networks on $\ERprob$ when the usage of ER topology
$\ERtoRING$ approaches zero.

>>>>

    \newcommand{\param}[1]{\texttt{#1}}
    \newcommand{\feedback}{\param{feedback}}
    \newcommand{\gain}{\param{gain}}
    \newcommand{\size}{N}
    \newcommand{\ERprob}{p}
    \newcommand{\RINGk}{k}
    \newcommand{\ERtoRING}{q}
    \newcommand{\FFlayers}{l}

@@

The genotypes' parameters $\theta\in\Theta$ are either integer or floating
point values and can be classified as affecting (1) the topology of
$\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
$\omega\Ix{fb}$. Optimized parameters of type (1) are the
occupation probability $\ERprob$ for sparse random topologies (Erdös-Renyi
topology), the neighborhood range $\RINGk$ for ring topologies, and a ratio
parameter $\ERtoRING$ used to interpolate between Erdös-Renyi and ring
topologies. Optimized parameters of type (2) are the feedback strength
$\feedback\propto \omega\Ix{fb}$ and the gain value
$\gain\propto\omega\Ix{rec}$.

@@

\begin{figure}
    {\footnotesize\input{pictures/freqrange}}
    \caption{
        The fitness values of four differently obtained genotypes when
        tested against deterministic challenges $C_\nu$ that yield simple
        sinusoidal waves with frequency $\nu$ for every seed value.  Here
        $G_0$ is an unoptimized genotype with parameters taken from
        \cite{FORCE}, whereas the genotypes $G\ix{high}$ and $G\ix{low}$
        did particularly well during the optimization for $C\ix{high}$ and
        $G\ix{low}$ depicted in FIG.  \ref{fig:freq_optimization}.  While
        these three genotypes had strict Erdös-Renyi topologies,
        $G\ix{high,ring}$ is a successful sample of an optimization for
        $C\ix{high}$ under an interpolation of Erdös-Renyi and ring topology
        (see FIG. \ref{fig:top_optimization} in the next section).
    }

        %$G\ix{high,ring}$ is
        %also optimized for $C\ix{high}$.

        %The covered frequency range for an unoptimized
        %genotype $G_0$, genotypes $G\ix{high}$ and $G\ix{high,ring}$ optimized for $C\ix{high}$ and $G\ix{low}$
        %that are optimized for $C\ix{high}$ respectively $C\ix{low}$, . Each
        %of the three genotypes was tested against constant challenges
        %$C_\nu$ that yielded simple sinusoidal waves with frequency
        %$\nu$ for every seed value. The topology of all genotypes was
        %a mere Erdös-Renyi topology.
        %The parameters for the unoptimized genotype $G_0$ were taken to be
        %default values from \cite{FORCE} (which also uses Erdös-Renyi
        %topologies), while we chose genotypes $G\ix{high}$ and $G\ix{low}$
        %that did particularly well when explicitly optimizing for high and
        %low frequencies (see FIG. \ref{fig:freq_optimization} for more
        %details).
    %}
    \label{fig:freqrange}
\end{figure}

\subsection{Topology Optimizations}
To find out whether the topology may be used to improve the learning for
high frequencies without increasing the feedback, we performed an
optimization in which the network topologies were an interpolation between
Erdös-Renyi and ring topology.
%the $\RINGk$ nearest neighbors are connected (for non integer $\RINGk$, the
%fractional part determines the probability to use $[\RINGk]$ or $[\RINGk]+1$).
For this simulation, the parameters $\gain = 1.5$ and $\feedback = 2$ are
kept constant during the optimization process, leaving the connection
probability for Erdös-Renyi topology $\ERprob$, the ring connection length $\RINGk$
and the interpolation value $\ERtoRING$ between Erdös-Renyi and ring as
free parameters.

The results of the optimization are depicted in FIG.
\ref{fig:top_optimization}. Unfortunately, the improvement in fitness after the
first few generations is extremely low. The development of the ration
parameter, however, unambiguously shows that the selection process prefers
generators with a high fraction of ring topology and with $\RINGk
\approx 1.5$. The Erdös-Renyi connection probability $\ERprob$ shows a very
broad distribution, which is due to the weak dependence of the generated
networks on $\ERprob$ when $\ERtoRING$ approaches zero.

The fitness curve of $G\ix{high,ring}$ in figure \ref{fig:freqrange}
confirms, that though $G\ix{high}$ is still better for high frequencies due
to the strong feedback, the ring topology provides an improvement compared
to the Erdös-Renyi topology when optimizing for $C\ix{high}$.
